---
title: "101-align-to-DSM"
author: "Mac Campbell"
date: "7/1/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(ggrepel)
```

## Using genomic reference?

```{r}
old<-read_tsv("outputs/longfin/old-counts-files.txt", col_names=c("File"))
old$Sample<-gsub("_sorted_flt.bam","",old$File)
oldreads<-read_tsv("outputs/longfin/old-counts.txt", col_names=c("Out"))
oldreads$Reads<-gsub(" + 0 mapped (100.00% : N/A)", "",oldreads$Out, fixed = TRUE)

oldtot<-bind_cols(old,oldreads) %>% select(-Out) %>% select(Sample, File, Reads) %>% mutate(Alignment="Original")
nrow(oldtot)
```

```{r}
new<-read_tsv("outputs/longfin/new-counts-files.txt", col_names=c("File"))
new$Sample<-gsub("_R1.sort.flt.bam","",new$File)
newreads<-read_tsv("outputs/longfin/new-counts.txt", col_names=c("Out"))
newreads$Reads<-gsub(" + 0 mapped (100.00% : N/A)", "",newreads$Out, fixed = TRUE)

newtot<-bind_cols(new,newreads) %>% select(-Out) %>% select(Sample, File, Reads) %>% mutate(Alignment="New Genome")
nrow(newtot)
```

```{r}
comb<-bind_rows(oldtot,newtot)
comb$Reads<-as.numeric(comb$Reads)
mean(as.numeric(oldtot$Reads))
mean(as.numeric(newtot$Reads))

#Hmmm… Basic idea

(578798.6-444797.9)/444797.9 * 100
write_tsv(comb, file="meta/counts.tsv")
## [1] 30.1262
```


30% more?

```{r}
ggplot(comb) +
  geom_histogram(aes(Reads, fill=Alignment), color="black", alpha=0.75, bins = 30) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust=0.5)) +
  scale_fill_viridis_d(option="D", direction=-1) +
  ylab("Count") 
```

### Basic Stuff

Make a PCA and see what it is about?

```{r}
sub<-comb %>% filter(Alignment=="New Genome") %>% filter(Reads > 1.5e5) %>%
  mutate(Path=paste0("/home/maccamp/data/longfin/",File))

mean(sub$Reads)
nrow(sub)
```

```{r}
write_tsv(select(sub, Path), col_names = FALSE, "bamlists/sub140.bamlist")
```

Generating beagle file

```{sh, eval=FALSE}
srun -p high -t 14:00:00 --mem=16G --nodes=1 $HOME/angsd/angsd -P 8 \
  -ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
  -bam $HOME/longfin/bamlists/sub140.bamlist -GL 1 \
  -doGLF 2 -doMajorMinor 1 -doMaf 2 -SNP_pval 1e-6 -minMapQ 10 -minQ 20 -minMaf 0.05 \
  -rf $HOME/delta-smelt/metadata/large-contigs.txt  -minInd 112 \
  -out $HOME/longfin/outputs/101/sub140-contigs > outputs/101/beagle-contigs.out 2> outputs/101/beagle-contigs.err &
```



-r lg01 produced 2028 snps
-rf HOME/delta-smelt/metadata/large-contigs.txt produced 66187 SNPs

Saglam et al found "13,402 high probability (P < 10−12) SNPs with a minor allele frequency over 0.05."

Getting a bit more info together

```{r}
meta<-sub %>% separate(Sample, into=c("Loc","Number"), sep="_", remove=FALSE)

#Factor N->S
meta$Location<-factor(meta$Loc, levels=c("YBAK","SKNA","PTLC","HRLC","FRAS","LWSH","COLR",
                                              "HUMB","SFBY","PETA","ALVS","SUIB","CHPI"))
```


#### PCA

```{sh, eval=FALSE}
python $HOME/pcangsd/pcangsd.py -beagle outputs/101/sub140.beagle.gz \
   -admix -o outputs/101/sub140-pca -threads 10

python $HOME/pcangsd/pcangsd.py -beagle outputs/101/sub140-contigs.beagle.gz \
   -admix -o outputs/101/sub140-contigs-pca -threads 10
```


```{r}
cov<-read_delim("outputs/101/sub140-contigs-pca.cov", col_names=FALSE, delim=" ") %>% as.matrix()
```

```{r}
#' @param samples character vector with the individuals IDs in the order in which
#' they were passed in the bamlist to angsd.
#' @param cov covariance matrix
covar2pcs <- function(samples, cov) {
  
  
  eig <- eigen(cov, symm = TRUE)
  PC <- as.data.frame(eig$vectors) %>%
    as_tibble() %>%
    setNames(sprintf("PC-%02d", 1:ncol(.)))
  
  samtib <- tibble(sample = samples)
  
  list(
    PCs = bind_cols(samtib, PC),
    eigevalues = eig$values
  )
}
```

```{r}
pca <- covar2pcs(meta$Sample, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

pp_meta <- pca_pairs %>%   # just keep the first 6 PCs around
  left_join(., meta, by = c("sample" = "Sample")) %>%
  mutate(group = Location) 
```

Plot    

```{r}
npc <- 3
pp_meta2 <- pp_meta %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) )

eig <- eigen(cov, symm = TRUE)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
```

```{r}
ggplot(pp_meta2, aes(x = val_x, y = val_y, color=Location)) +
  geom_point() +
  facet_grid(PCx ~ PCy)
```

```{r}
sub12<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-02") )
text<- sub12 %>% select(sample,Location,val_x,val_y) %>% group_by(Location) %>%
  mutate(meanX=mean(val_x), meanY=mean(val_y)) %>% select(Location, meanX, meanY) %>%
  unique()   

ggplot(sub12, aes(x = val_x, y = val_y, color=Location))+
  geom_point(size = 2, alpha=0.5) +
  geom_text_repel(data=text, aes(x=meanX, y=meanY, label=Location), max.overlaps = Inf) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = ""))+
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis_d(option="C")

ggsave("outputs/101/lfs-dsm-reference-pc12.jpeg")
```

```{r}
pc12<-ggplot(sub12, aes(x = val_x, y = val_y, color=Location))+
  geom_point(size = 2, alpha=0.75) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = ""))+
  theme(legend.position = "")+
  theme(plot.title = element_text(hjust = 0.5)) 

sub13<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-03") )

pc13<-ggplot(sub13, aes(x = val_x, y = val_y, color = Location)) +
  geom_point(size = 2, alpha=0.75) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = ""))+
  theme(legend.position = "")+
  theme(plot.title = element_text(hjust = 0.5)) 

pcs<-ggarrange(pc12, pc13, ncol = 2)
pcs
```

```{sh, eval=FALSE}
$HOME/angsd/misc/NGSadmix  -likes outputs/101/sub140-contigs.beagle.gz -K 2 -o outputs/101/sub140-k2 -P 3
$HOME/angsd/misc/NGSadmix  -likes outputs/101/sub140-contigs.beagle.gz -K 3 -o outputs/101/sub140-k3 -P 3

```

```{r}
q2<-read_delim("outputs/101/sub140-k2.qopt", delim=" ", col_names=FALSE) %>% select(X1, X2) %>%
  rename(Q1=X1, Q2=X2)

df2<-bind_cols(q2, meta)            

q2s<-df2 %>% select(Location, Sample, Q1, Q2) %>% arrange(Location,Sample) %>%
  mutate(Index=1:n()) %>% gather(key=Ancestry, value=Q, 3:4) %>% mutate(K=2)

pops2<-q2s  %>% group_by(Location) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  select(Location,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2)) %>% ungroup() %>% unique()

```

```{r}
p2 <- ggplot(q2s) + 
  geom_col(aes(x=Index,y=Q, fill=Ancestry), color="NA", size = 0, width = 1) +
  geom_segment(data=pops2, x = pops2$Start - 0.5, y=0, xend = pops2$Start-0.5, yend=1, alpha=0.9, size=0.25) +
  geom_segment(data=pops2, x = pops2$Stop[length(pops2$Stop)]  + 
                 0.5, y=0, xend= pops2$Stop[length(pops2$Stop)] + 0.5, yend=1,  alpha=0.9, size=0.25) +
  geom_segment(x=0.5, xend= pops2$Stop[length(pops2$Stop)]+0.5, y=1, yend=1, alpha=0.9, size=0.25) +
  geom_segment(x=0.5, xend= pops2$Stop[length(pops2$Stop)]+0.5, y=0, yend=0, alpha=0.9, size=0.25) +
  ylim(-0.1,1.01) +
  xlim(-0.1, pops2$Stop[length(pops2$Stop)]+1) +
  theme(panel.background = element_blank()) +
  xlab("") + theme(legend.position = "") +
 # theme(axis.text.x = element_blank(), axis.ticks.x=element_blank()) +
  ggtitle("K = 2") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_x_continuous(breaks=pops2$Position, labels=pops2$Location) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  scale_fill_viridis_d(option="B")


p2

ggsave("outputs/101/lfs-dsm-reference-k2.jpg")
```


```{r}
q3<-read_delim("outputs/101/sub140-k3.qopt", delim=" ", col_names=FALSE) %>% select(X1, X2, X3) %>%
  rename(Q1=X1, Q2=X2, Q3=X3)

df3<-bind_cols(q3, meta)            

q3s<-df3 %>% select(Location, Sample, Q1, Q2, Q3) %>% arrange(Location,Sample) %>%
  mutate(Index=1:n()) %>% gather(key=Ancestry, value=Q, 3:5) %>% mutate(K=3)

pops3<-q3s  %>% group_by(Location) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  select(Location,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2)) %>% ungroup() %>% unique()

```

```{r}
p3 <- ggplot(q3s) + 
  geom_col(aes(x=Index,y=Q, fill=Ancestry), color="NA", size = 0, width = 1) +
  geom_segment(data=pops2, x = pops2$Start - 0.5, y=0, xend = pops2$Start-0.5, yend=1, alpha=0.9, size=0.25) +
  geom_segment(data=pops2, x = pops2$Stop[length(pops2$Stop)]  + 
                 0.5, y=0, xend= pops2$Stop[length(pops2$Stop)] + 0.5, yend=1,  alpha=0.9, size=0.25) +
  geom_segment(x=0.5, xend= pops2$Stop[length(pops2$Stop)]+0.5, y=1, yend=1, alpha=0.9, size=0.25) +
  geom_segment(x=0.5, xend= pops2$Stop[length(pops2$Stop)]+0.5, y=0, yend=0, alpha=0.9, size=0.25) +
  ylim(-0.1,1.01) +
  xlim(-0.1, pops2$Stop[length(pops2$Stop)]+1) +
  theme(panel.background = element_blank()) +
  xlab("") + theme(legend.position = "") +
 # theme(axis.text.x = element_blank(), axis.ticks.x=element_blank()) +
  ggtitle("K = 3") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_x_continuous(breaks=pops2$Position, labels=pops2$Location) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  scale_fill_viridis_d(option="B")


p3

ggsave("outputs/101/lfs-dsm-reference-k3.jpg")
```
