---
title: "104-report"
author: "Mac Campbell"
date: "2022-12-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ggrepel)
library(ggpubr)
library(adegenet)
library(vcfR)
library(ggtree)
library(phytools)
```
     
Some ideas:

Reads available after aligning to DSM genome and de novo assembly.

Reads available as function of DNA amount.

# Where do the new samples end up


```{r}
lfs<-read_csv("meta/175-sub-meta.csv") %>% separate(Sample, into=c("Pre","Num1","Num2"), remove = FALSE)

new<-lfs %>% filter(Pre %in% c("19","20","Nksk")) %>% mutate(Collection=paste0(Pre,"-",Num1)) %>% select(-Location) %>%
  relocate(Collection) %>%
  mutate(Location=ifelse(Collection %in% c("19-0222","19-0225"),"EURS",
                  ifelse(Collection %in% c("20-1325","20-1329","20-1330","20-1373","20-1375"),"EELR",
                  ifelse(Collection %in% c("20-1291"), "KLAMR",
                  ifelse(Collection %in% c("20-1356"), "MADR",
                  ifelse(Collection %in% c("Nksk-2018"),"NOOK","None"))))))

comb<-lfs %>% filter(!Pre %in% c("19","20","Nksk")) %>% bind_rows(new)

```

```{r}
cov<-read_delim("outputs/102/175-sub50-pca.cov", col_names=FALSE, delim=" ") %>% as.matrix()
meta<-comb
meta$Location<-factor(meta$Location, levels=c("YBAK","SKNA","PTLC","HRLC","FRAS","NOOK","LWSH","COLR",
                                             "KLAMR","EURS","EELR","MADR","HUMB","SFBY","PETA","ALVS","SUIB","CHPI"))
```

```{r}
pca <- covar2pcs(meta$Sample, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

pp_meta <- pca_pairs %>%   # just keep the first 6 PCs around
  left_join(., meta, by = c("sample" = "Sample")) %>%
  mutate(group = Location) 

npc <- 3
pp_meta2 <- pp_meta %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) )

eig <- eigen(cov, symm = TRUE)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
```

```{r}
ggplot(pp_meta2, aes(x = val_x, y = val_y, color=Location)) +
  geom_point() +
  facet_grid(PCx ~ PCy)
```

```{r}
sub12<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-02") )
text<- sub12 %>% select(sample,Location,val_x,val_y) %>% group_by(Location) %>%
  mutate(meanX=mean(val_x), meanY=mean(val_y)) %>% select(Location, meanX, meanY) %>%
  unique()   

pc12<-ggplot(sub12, aes(x = val_x, y = val_y, color=Location))+
  geom_point(size = 2, alpha=0.5) +
  geom_label_repel(data=text, aes(x=meanX, y=meanY, label=Location), max.overlaps = Inf) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = ""))+
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis_d(option="H") +
  theme(legend.position = "")

sub13<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-03") )
text<- sub13 %>% select(sample,Location,val_x,val_y) %>% group_by(Location) %>%
  mutate(meanX=mean(val_x), meanY=mean(val_y)) %>% select(Location, meanX, meanY) %>%
  unique()   

pc13<-ggplot(sub13, aes(x = val_x, y = val_y, color=Location))+
  geom_point(size = 2, alpha=0.5) +
  geom_label_repel(data=text, aes(x=meanX, y=meanY, label=Location), max.overlaps = Inf) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = ""))+
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_viridis_d(option="H") +
  theme(legend.position = "")
```

```{r}
ggarrange(pc12, pc13)

ggsave("outputs/104/longfin175-pc123.jpeg", width=10, height=5)
```
```{r}
meta %>% group_by(Location) %>% summarize(Count=n(), AvgReads=mean(Reads)) %>% arrange(AvgReads)
```

```{r}
new<-meta %>% filter(Location %in% c("KLAMR","EURS","EELR","MADR","NOOK"))
ggplot(new) +
  geom_histogram(aes(x=Reads, fill=Collection))

new %>% filter(Location=="NOOK") %>% select(Reads)
```

## IBS version

```{sh, eval=FALSE}
srun -p high -t 14:00:00 --mem=16G --nodes=1 $HOME/angsd/angsd -P 36 \
  -ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
  -bam $HOME/longfin/bamlists/175-sub.bamlist -GL 1 \
  -doGLF 2 -doMajorMinor 1 -doMaf 2 -SNP_pval 1e-3 -minMapQ 10 -minQ 20 -minMaf 0.05 \
  -rf $HOME/delta-smelt/metadata/large-contigs.txt  -minInd 88 \
  -out $HOME/longfin/outputs/102/175-sub50> outputs/102/beagle50.out 2> outputs/102/beagle50.err &
  
```


```{r}
m <- as.matrix(read.table("outputs/102/175-ibs-gl.covMat"))
meta<-meta
eig <- eigen(m)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
covs<-eig$vectors[,1:3] %>% as_tibble() %>% bind_cols(meta)

text12<-covs %>% select(Sample, Location, V1, V2) %>%
  group_by(Location) %>% summarize(Count=n(), x=mean(V1), y=mean(V2))

covs12<-ggplot(covs) +
  geom_point(aes(x=V1, y=V2, fill=Location), pch=21, alpha=0.75) +
  geom_text_repel(data=text12, aes(x=x, y=y, label=Location), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = "")) +
  theme_bw() +
  theme(legend.position="")
covs12
```

```{r}
ggplot(covs%>% filter(Location %in% c("EELR","EURS","NOOK"))) +
  geom_point(aes(x=V1, y=V2, fill=Reads), pch=21, alpha=0.75) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = "")) +
  facet_wrap(.~Location) +
  theme_bw()
```
```{r}

text13<-covs %>% select(Sample, Location, V1, V3) %>%
  group_by(Location) %>% summarize(Count=n(), x=mean(V1), y=mean(V3))

covs13<-ggplot(covs) +
  geom_point(aes(x=V1, y=V3, fill=Location), pch=21, alpha=0.75) +
  geom_text_repel(data=text13, aes(x=x, y=y, label=Location), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = "")) +
  theme_bw()
covs13
``` 

    
```{r}
ggarrange(covs12, covs13, widths=c(1,1.3))
ggsave("outputs/104/longfin175-ibs-pc123.jpeg", width=10, height=5)

```
## Read Counts


```{r}
existing<-read_tsv("meta/counts.tsv") %>% filter(Alignment !="Original")
existing<-existing %>% filter(Alignment=="New Genome") %>% filter(Reads > 1.5e5) %>%
  mutate(Path=paste0("/home/maccamp/data/longfin/",File))

ggplot(existing) + geom_histogram(aes(x=Reads))
```

```{r}
new<-read_csv("outputs/102/newseq.dat", col_names = c("Sample","Sort","Dedup"))

ggplot(new) + geom_histogram(aes(x=Dedup))
```

```{r}
merge<-existing %>% select(Sample, Reads) %>% mutate(Origin="Existing") %>%
  bind_rows(select(new, Sample, Dedup) %>% rename(Reads=Dedup) %>% mutate(Origin="New Sequencing"))

ggplot(merge, aes(fill=Origin)) + geom_histogram(aes(x=Reads)) +
  theme_bw() +
  scale_fill_viridis_d(option="H") +
  facet_wrap(.~Origin, ncol=1) +
  ylab("Count")

ggsave("outputs/104/longfin-counts.jpeg", width=6, height=4)

``` 

```{r}
quants<-read_csv("meta/original-quants.csv") %>% mutate(DNA="Starting")
conc<-read_csv("meta/concentrated-quants.csv") %>% rename(Samples=SampleID,`conc ng/ul` = `Conc ng/ul`) %>% mutate(DNA="Concentrated")
quants<-bind_rows(quants, conc)
quants

quants$Samples<-gsub(" ","-",quants$Samples)

quants<-new %>% left_join(quants, by=c("Sample"="Samples"))

ggplot(quants %>% filter(DNA=="Concentrated"), aes(x=`conc ng/ul`, y=Dedup)) +
  geom_point(pch=21, alpha=0.75, color="black", fill="black") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  ylab("Reads Remaining for Analyses\n") +
  xlab("\nConcentration of DNA in ng/ul") +
  geom_hline(aes(yintercept=1.5e5), lty=2)

ggsave("outputs/104/dna-reads.jpeg", width=6, height=4)
```

Droop out nooksack     

```{r}

ggplot(quants[!startsWith(quants$Sample, "Nksk"),] %>% filter(DNA=="Concentrated"), aes(x=`conc ng/ul`, y=Dedup)) +
  geom_point(pch=21, alpha=0.75, color="black", fill="black") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  ylab("Reads Remaining for Analyses\n") +
  xlab("\nConcentration of DNA in ng/ul") +
  geom_hline(aes(yintercept=1.5e5), lty=2)

ggsave("outputs/104/dna-reads-no-nooksack.jpeg", width=6, height=4)
```


## What about those other species of smelts??

~/data/forage-fishes symlinking some to ~/longfin/data/forage-fishes    
EULN_006, 007, 010     
WAKS 010, 011, 012    
14 DSM (can downsample)

There is no mtDNA in reference, dang.  How to figure out mystery fish from phylogeny? Blast to mtDNA, make new reference for mtDNA???

No H. transpacificus, no A. elongatus, No S. starksi, No S. thaleichthys (S. lanceolatus) exists.   


Align reference EULN WAKS DSM

```{sh, eval=FALSE}
# in data/forage-fish
ls | grep R1 | perl -pe 's/.fastq//g' > forward
ls | grep R2 | perl -pe 's/.fastq//g' > reverse
ls | grep R1 | perl -pe 's/_R1.fastq//g' > name
paste forward reverse name  > samples.txt
bash $HOME/missouri-trout/doAlign-unzipped.sh samples.txt /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa 
```    

(Can include some Longfin)
ALVS_097_R1.sort.flt.bam 
ALVS_098_R1.sort.flt.bam
CHPI_707_R1.sort.flt.bam
HUMB_004_R1.sort.flt.bam
HUMB_005_R1.sort.flt.bam
HUMB_006_R1.sort.flt.bam
SUIB_033_R1.sort.flt.bam
SUIB_044_R1.sort.flt.bam
YBAK_003_R1.sort.flt.bam
YBAK_007_R1.sort.flt.bam
COLR_030_R1.sort.flt.bam

```{r}
refs<-read_csv("outputs/102/forage-fishes.dat", col_names = c("Sample","Sort","Dedup")) %>% 
  mutate(Path=paste0("data/forage-fishes/",Sample,".sort.flt.bam"))

ggplot(refs) +
  geom_histogram(aes(x=Dedup))
```

Mystery fish:
(base) maccamp@farm:~/longfin/data$ cat ./alt-1/alt-1.dat ./alt-2/alt-2.dat ./alt-3/alt-3.dat > alt.dat

```{r}
dat<-read_csv("outputs/104/alt.dat", col_names = c("Sample","Sort","Dedup"))

ggplot(dat) +
  geom_histogram(aes(x=Dedup))
```

Making calls with phylogeny, want quite a few reads
```{r}
ds<-dat %>% filter(Dedup > 5e5) %>% mutate(Path=paste0("data/alt-bams/",Sample,".sort.flt.bam"))
nrow(ds)
```


```{r}
combine<-bind_rows(refs,ds)
write_csv(combine,"outputs/102/combine.csv")
```

```{r}
m216<-read_csv("meta/216.csv")
write_tsv(m216 %>% select(Path), col_names = FALSE, file="bamlists/216.bamlist")
write_tsv(m216 %>% select(Sample), col_names = FALSE, file="bamlists/216.samples")

```



```{sh, eval=FALSE}
srun -p high -t 18:00:00 --mem=16G --nodes=1 angsd -P 24 \
-bam bamlists/216.bamlist \
-out /home/maccamp/longfin/outputs/104/plink \
-anc /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-rf $HOME/delta-smelt/metadata/large-contigs.txt  \
-minInd 195 -minMaf 0.05  -minMapQ 10 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 4 -doPost 1 -postCutoff 0.95 -doPlink 2 >outputs/104/std.out 2>outputs/104/std.err &
```     

	-> Number of sites retained after filtering: 100035 

## Process


```{sh, eval=FALSE}
plink --tped plink.tped --tfam plink.tfam  --out binary --recode --allow-extra-chr --noweb
plink --ped binary.ped --map binary.map --recode vcf --allow-extra-chr -out recode
bcftools +prune -l 0.25 -w 10000 recode.vcf  -Ov -o recode.prune.vcf
bcftools reheader --samples ../../bamlists/216.samples -o recode.prune.reheadered.vcf recode.prune.vcf

```

After pruning, 7380 snps

onvert to phylip, then nexus. Create a missing data set

```{sh, eval=FALSE}
source activate py2; ~/github/mccloud-rrt/vcf2phylip.py -i recode.prune.reheadered.vcf; conda deactivate;
seqConverter.pl -drecode.prune.reheadered.min4.phy -on

#Corrected for ascertainment bias
conda activate py3; ~/github/mccloud-rrt/103-remove-invariant.py -p recode.prune.reheadered.min4.phy -o recode.prune.reheadered.min4.asc.phy; conda deactivate

seqConverter.pl -drecode.prune.reheadered.min4.asc.phy -on

#Now remove missing?? 
 ~/github/relict-dace/201.1-removeMissing.pl recode.prune.reheadered.min4.asc.phy  > filtered.phy
seqConverter.pl -dfiltered.phy -on

Our dumped samples are DELS, WAKS and EULN. 
```

Calculate allele freqs:

recode.prune.reheadered.vcf

```{sh, eval=FALSE}
bcftools view -S outputs/104/lfs.tsv outputs/104/recode.prune.reheadered.vcf > outputs/104/lfs.vcf
bcftools view -S outputs/104/non-lfs.tsv outputs/104/recode.prune.reheadered.vcf > outputs/104/non-lfs.vcf
bcftools view -S outputs/104/delta.tsv outputs/104/recode.prune.reheadered.vcf > outputs/104/delta.vcf


vcftools --vcf outputs/104/lfs.vcf  --freq --out outputs/104/lfs
vcftools --vcf outputs/104/non-lfs.vcf --freq --out outputs/104/non-lfs
vcftools --vcf outputs/104/delta.vcf --freq --out outputs/104/delta

```

141 lfs / 7380 sites
72 non-lfs / 7380 sites
13 dsm    
```{r}
freq1<-read_tsv("outputs/104/lfs.frq", skip = 1, col_names = FALSE) %>% 
  rename(Chromosome=X1, Position=X2, NAllelesWinter=X3, NLFS=X4, Major=X5, Minor=X6) %>%
  separate(Major, into=c("Major","MajorFrequencyLFS"), sep=":") %>% 
  separate(Minor, into=c("Minor","MinorFrequencyLFS"), sep=":")
freq1$MajorFrequencyLFS<-as.numeric(freq1$MajorFrequencyLFS)
freq1$MinorFrequencyLFS<-as.numeric(freq1$MinorFrequencyLFS)

freq2<-read_tsv("outputs/104/delta.frq", skip = 1, col_names = FALSE) %>% 
  rename(Chromosome=X1, Position=X2, NAlleles=X3, N=X4, Major=X5, Minor=X6)%>%
  separate(Major, into=c("Major","MajorFrequency"), sep=":") %>% separate(Minor, into=c("Minor","MinorFrequency"), sep=":")
freq2$MajorFrequency<-as.numeric(freq2$MajorFrequency)
freq2$MinorFrequency<-as.numeric(freq2$MinorFrequency)

comb<-left_join(freq2, freq1)

merge<-comb %>% mutate(AbsoluteDifference=abs(MajorFrequencyLFS-MajorFrequency)) %>% 
  select(Chromosome, Position, Major, Minor, MajorFrequency, MinorFrequency, MajorFrequencyLFS, MinorFrequencyLFS, AbsoluteDifference, N, NLFS) %>% 
  filter(N > 12*2, NLFS > 113*2) %>% arrange(-AbsoluteDifference)

write_csv(merge,"outputs/104/LFS-non-LFS-allele-freqs.csv")

```

```{r}
merge %>% filter(MajorFrequencyLFS > 0.95) %>% filter (AbsoluteDifference > 0.80)
```
Local test:

```{sh, eval=FALSE}
seqConverter.pl -drecode.prune.reheadered.min4.asc.phy -of
selectSites.pl -s 1-1000 recode.prune.reheadered.min4.asc.fasta  > sub.fasta
seqConverter.pl -dsub.fasta -ope
iqtree -s sub.phylip -st DNA -m GTR+F+ASC -bb 1000 -alrt 1000 -redo

```

Removed some low quality samples from dumpers (DELS 710 733, an ALVS sample, now 213 samples)
iqtree -s align.phy -st DNA -m GTR+F+ASC -bb 1000 -alrt 1000 -redo
iqtree -s align.phy.varsites.phy -st DNA -m GTR+F+ASC -bb 1000 -alrt 1000 -redo

-T 2 (two cores available apparently)    

```{sh, eval=FALSE}
srun -p high -t 5:00:00 --nodes=1 iqtree -s align.phy.varsites.phy -st DNA -m MFP+ASC -bb 1000 -alrt 1000 -T AUTO -redo
```

## Get samples by species

```{r}
snps<-read.vcfR("outputs/104/recode.prune.reheadered.vcf")
gen<-vcfR2genind(snps)
```

```{r}
x<- tab(gen, NA.method="mean")
pca <- dudi.pca(x,scannf=FALSE,scale=FALSE,nf=5)
plot(pca$li)
```


```{r}
d<- pca$l1 %>% as_tibble()
d$Sample<-rownames(pca$l1)

#Pull out cumulative variance
eig<-pca$eig
sum<-sum(eig)
var<-(eig/sum)*100

d<- d %>% left_join(m216)

#text<-dsba %>% group_by(`River Basin`) %>% mutate(meanx=mean(RS1), meany=mean(RS2)) %>% 
 # select(`River Basin`, meanx, meany) %>% unique()

pc12<-ggplot(d) +
  geom_point(aes(x=RS1, y=RS2), pch=21, alpha=0.75, fill="black") +
#  geom_label_repel(data=text, aes(x=meanx, y=meany, label=`River Basin`),
  #                 fill="white", max.overlaps = Inf, alpha=0.9) +
  #scale_fill_viridis_d(option="H") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "") +
  xlab(paste0("PC1"," ",round(var[1],2),"%")) +
  ylab(paste0("PC2"," ",round(var[2],2),"%")) 

pc12
```    
```{r}
longs<-d %>% filter(RS1 >0)
other<-d %>% filter(RS1 <0) %>% filter(!(Sample %in% c("DELS_710","DELS_733")))
```
Removed some low quality samples from dumpers (DELS 710 733, an ALVS sample, now 213 samples)     

```{r}
ggplot(other, aes(x=RS1, y=RS2)) +
  geom_point(pch=21, alpha=0.5, fill="black") + geom_text_repel(aes(label=Sample), max.overlaps = Inf)
ggsave("outputs/104/smelts.pdf", width=22, height=22)
```   

Create a bamllist of reference samples

```{r}
ls<-c("ALVS_098_R1","CHPI_707_R1","HUMB_004_R1","HUMB_005_R1","HUMB_006_R1","SUIB_033_R1","SUIB_044_R1","YBAK_003_R1","YBAK_007_R1","COLR_030_R1") %>%
  as_tibble() %>% rename(Sample=value)

smelts<-select(other, Sample) %>% bind_rows(ls)
m72<-smelts %>% left_join(m216)
write_tsv(m72 %>% select(Path), col_names=FALSE, file="bamlists/72.bamlist")
write_tsv(m72 %>% select(Sample), col_names=FALSE, file="bamlists/72.samples")
```

```{sh, eval=FALSE}
srun -p high -t 18:00:00 --mem=16G --nodes=1 angsd -P 24 \
-bam bamlists/72.bamlist \
-out /home/maccamp/longfin/outputs/104/refs/plink \
-anc /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-rf $HOME/delta-smelt/metadata/large-contigs.txt  \
-minInd 65 -minMaf 0.05  -minMapQ 10 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 4 -doPost 1 -postCutoff 0.95 -doPlink 2 >outputs/104/refs/std.out 2>outputs/104/refs/std.err &
```

	-> Number of sites retained after filtering: 103435 

in outputs/104/refs/ 

```{sh, eval=FALSE}
plink --tped plink.tped --tfam plink.tfam  --out binary --recode --allow-extra-chr --noweb
plink --ped binary.ped --map binary.map --recode vcf --allow-extra-chr -out recode
bcftools +prune -l 0.25 -w 10000 recode.vcf  -Ov -o recode.prune.vcf
bcftools reheader --samples ../../../bamlists/72.samples -o recode.prune.reheadered.vcf recode.prune.vcf
```

```{sh, eval=FALSE}
source activate py2; ~/github/mccloud-rrt/vcf2phylip.py -i recode.prune.reheadered.vcf; conda deactivate;
seqConverter.pl -drecode.prune.reheadered.min4.phy -on

#Corrected for ascertainment bias
conda activate py3; ~/github/mccloud-rrt/103-remove-invariant.py -p recode.prune.reheadered.min4.phy -o recode.prune.reheadered.min4.asc.phy; conda deactivate

seqConverter.pl -drecode.prune.reheadered.min4.asc.phy -on
#1016 SNPs

#Now remove missing?? 
 ~/github/relict-dace/201.1-removeMissing.pl recode.prune.reheadered.min4.asc.phy  > filtered.phy
seqConverter.pl -dfiltered.phy -on
```

Dumps eulachon and some other samples.  

```{sh, eval=FALSE}
iqtree -T AUTO -s recode.prune.reheadered.min4.asc.phy -st DNA -m GTR+F+ASC -bb 1000 -alrt 1000 -redo
``` 


CHPI_707_R1 is oddly placed, but otherwise, looks good. CHPI_707_R1 is not oddly placed in ML phylogeny.    

```{r}
snps<-read.vcfR("outputs/104/refs/recode.prune.reheadered.vcf")
gen<-vcfR2genind(snps)
x<- tab(gen, NA.method="mean")
pca <- dudi.pca(x,scannf=FALSE,scale=FALSE,nf=5)
plot(pca$li)
```


```{r}
d<- pca$l1 %>% as_tibble()
d$Sample<-rownames(pca$l1)

#Pull out cumulative variance
eig<-pca$eig
sum<-sum(eig)
var<-(eig/sum)*100

d<- d %>% left_join(m216)

#text<-dsba %>% group_by(`River Basin`) %>% mutate(meanx=mean(RS1), meany=mean(RS2)) %>% 
 # select(`River Basin`, meanx, meany) %>% unique()

pc12<-ggplot(d) +
  geom_point(aes(x=RS1, y=RS2), pch=21, alpha=0.75, fill="black") +
#  geom_label_repel(data=text, aes(x=meanx, y=meany, label=`River Basin`),
  #                 fill="white", max.overlaps = Inf, alpha=0.9) +
  #scale_fill_viridis_d(option="H") +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "") +
  xlab(paste0("PC1"," ",round(var[1],2),"%")) +
  ylab(paste0("PC2"," ",round(var[2],2),"%")) +
  geom_text_repel(aes(x=RS1, y=RS2, label=Sample), max.overlaps = Inf)
  ggsave("outputs/104/ref-smelts.pdf", width=22, height=22)

pc12
``` 


```{r}
waks<-d %>% filter(RS1 < 0, RS2 < -1 ) %>% mutate(Species="Wakasagi")
dels<-d %>% filter(RS1 < 0, RS2 > 0)  %>% filter(RS5 < 0) %>% mutate(Species="Delta Smelt")
surf<-d %>% filter(RS1 < 0, RS2 > 0)  %>% filter(RS5 > 0) %>% mutate(Species="Surf Smelt")
euln<-d %>% filter(RS1 > 0, RS2 > 0.35) %>% mutate(Species="Eulachon")
nigh<-d %>% filter(RS1 > 0, RS2 < 0.35) %>% filter(RS4 < 0) %>%  mutate(Species="Night Smelt") %>% filter(!(Sample %in% ls$Sample))
long<-d %>% filter(RS1 > 0, RS2 < 0.35) %>% filter(RS4 < 0) %>%  mutate(Species="Longfin Smelt") %>% filter((Sample %in% ls$Sample))
whit<-d %>% filter(RS1 > 0, RS2 < 0.35) %>% filter(RS4 > 0)%>%  mutate(Species="Whitebait Smelt") %>% filter(!(Sample %in% ls$Sample))
```

```{r}
names<-bind_rows(waks,dels,surf,euln,nigh,long,whit) %>% select(Sample,Species)
names
write_tsv(names,"meta/non-lfs.tsv")
write_tsv(names %>% select(Sample), col_names = FALSE,"outputs/104/non-lfs.tsv")

write_tsv(names %>% filter (Species=="Delta Smelt") %>% select(Sample), col_names = FALSE,"outputs/104/delta.tsv")

```
Nooks
```{r}
nooks<-read_csv("meta/meta260.csv") %>% filter(Loc == "NOOK")
nooks$Sample<-gsub("plate-1","HS2253_AGTCAA",nooks$Sample)
nooks
```


```{r}
t<-read.tree("outputs/104/align.phy.varsites.phy.contree")
u<-midpoint.root(t) 
u<-as.polytomy(u, feature='node.label', fun=function(x) as.numeric(x) < 75)
v<-ggtree(u) %<+% m216 
data<-v$data
ddf<-data  %>% left_join(names, by=c("label"="Sample")) %>% filter(isTip==TRUE) %>% select(label, Species) %>%  replace(is.na(.), "Larval Smelt") %>% mutate(Category=ifelse(label %in% nooks$Sample,"Nooksack",Species))
ddf$Species<-ddf$Category
ddf$Species<-factor(ddf$Species, levels=c("Larval Smelt","Nooksack", "Longfin Smelt", "Night Smelt", "Eulachon", "Whitebait Smelt", "Wakasagi", "Surf Smelt", "Delta Smelt" ))
v$data<-v$data %>% left_join(ddf)
```

```{r}
ddf %>% filter(!label %in% names$Sample) %>% select(label) %>% write_tsv(col_names = FALSE, file="outputs/104/lfs.tsv")
```

```{r}
df<-ggtree(u)$data %>% filter(isTip==FALSE)
df$label<-as.numeric(df$label)
df<-df %>% mutate(Support=ifelse(label < 85, ">= 75 < 85",
                             ifelse(label >= 85 & label < 95, ">= 85 < 95", 
                             ifelse(label >= 95, ">= 95","NA")))) %>%
  na.omit()
```

```{r}
v +
  geom_point(data=df, aes(x=x, y=y, color=Support), pch=18, cex=4, alpha=1) +
  geom_tippoint(aes(fill=Species), pch=21, alpha=1, cex=4, offset=0.01) +
  geom_treescale() +
  xlim(0, 0.1) +
  scale_fill_viridis_d(option = "H") +
  scale_color_manual(values=c("grey50","grey","black"))
  
ggsave("outputs/104/lfs-tree.pdf", width=8.5, height=11)
```

```{r}
ddf %>% group_by(Species) %>% summarize(Count=n())
```



## how many samples/plate

```{r}
plate1<-dat[startsWith(dat$Sample, "HS2253"),] 
plate1<-plate1 %>% mutate(Plate="plate-1")
plate1$Barcode<-gsub("HS2253_AGTCAA-","",plate1$Sample)

plate23<-dat[(!startsWith(dat$Sample, "HS2253")),] 
plate23<-plate23 %>% separate(Sample, into = c("p1","p2","Barcode"), remove=FALSE) %>% mutate(Plate=paste0(p1,"-",p2)) %>% select(-p1,-p2)

plates<-bind_rows(plate1, plate23)


```

```{r}
plates %>% filter(Sort > 5e3) %>% group_by(Plate) %>% summarize(Count=n())
```

Expected plate3 to be 1/2 a plate.

## Pool longfin and then do some popgen

```{r}
newseq<-ddf %>% filter(Species == "Larval Smelt") %>% left_join(ds, by=c("label"="Sample"))
write_tsv(select(newseq, Path), col_names = FALSE, file="bamlists/141.bamlists")
```

141 samples. Presumably Nooksack is different from all these guys.

```{sh, eval=FALSE}
srun -p high -t 14:00:00 --mem=16G --nodes=1 $HOME/angsd/angsd -P 36 \
  -ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
  -bam $HOME/longfin/bamlists/141.bamlists -GL 1 \
  -doGLF 2 -doMajorMinor 1 -doMaf 2 -SNP_pval 1e-3 -minMapQ 10 -minQ 20 -minMaf 0.05 \
  -rf $HOME/delta-smelt/metadata/large-contigs.txt  -minInd 71 \
  -out $HOME/longfin/outputs/104/141> outputs/104/beagle-141.out 2> outputs/104/beagle-141.err &
  
srun -p high -t 12:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 36  \
-bam bamlists/141.bamlists -rf $HOME/delta-smelt/metadata/large-contigs.txt \
-ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-minInd 71 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-3 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/104/141-ibs > outputs/104/141-ibs.stdout 2> outputs/104/141-ibs.stderr &

  
srun -p high -t 14:00:00 --mem=16G --nodes=1 $HOME/angsd/angsd -P 36 \
  -ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
  -bam $HOME/longfin/bamlists/141.bamlists -GL 1 \
  -doGLF 2 -doMajorMinor 1 -doMaf 2 -SNP_pval 1e-3 -minMapQ 10 -minQ 20 -minMaf 0.05 \
  -r lg01  -minInd 71 \
  -out $HOME/longfin/outputs/104/141-sub > outputs/104/beagle-sub-141.out 2> outputs/104/beagle-sub-141.err &


srun -p high -t 12:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 36  \
-bam bamlists/141.bamlists -r lg01 \
-ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-minInd 71 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-3 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/104/141-lg01-ibs > outputs/104/141-lg01-ibs.stdout 2> outputs/104/141-lg01-ibs.stderr &
```

4143 sites on lg01 w/o ibs
4518 sites on lg01 w/ibs

```{r}
m <- as.matrix(read.table("outputs/104/141-ibs.covMat"))
meta<-newseq
eig <- eigen(m)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
covs<-eig$vectors[,1:3] %>% as_tibble() %>% bind_cols(meta)

#text12<-covs %>% select(Sample, Location, V1, V2) %>%
 # group_by(Location) %>% summarize(Count=n(), x=mean(V1), y=mean(V2))

covs12<-ggplot(covs) +
    geom_point(aes(x=V1, y=V2), pch=21, fill="black", alpha=0.75) +
#  geom_point(aes(x=V1, y=V2, fill=Location), pch=21, alpha=0.75) +
#  geom_text_repel(data=text12, aes(x=x, y=y, label=Location), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = "")) +
  theme_bw() +
  theme(legend.position="")
covs12
```

```{r}

#text13<-covs %>% select(Sample, Location, V1, V3) %>%
 # group_by(Location) %>% summarize(Count=n(), x=mean(V1), y=mean(V3))

covs13<-ggplot(covs) +
  geom_point(aes(x=V1, y=V3), pch=21, fill="black",alpha=0.75) +
  #  geom_point(aes(x=V1, y=V3, fill=Location), pch=21, alpha=0.75) +
#  geom_text_repel(data=text13, aes(x=x, y=y, label=Location), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = "")) +
  theme_bw()
covs13
``` 

```{r}
mynooks<-covs %>% filter(V1 > 0.1)
mynooks
```

```{r}
v2<-v
names2<-bind_rows(names, mynooks %>% select(label) %>% rename(Sample=label) %>% mutate(Species="Longfin Smelt"))
data2<-v2$data %>% select(-Species)

v2ddf<- data2  %>% left_join(names2, by=c("label"="Sample")) %>% filter(isTip==TRUE) %>% select(label, Species) %>%  replace(is.na(.), "Larval Smelt")
v2ddf$Species<-factor(v2ddf$Species, levels=c("Larval Smelt", "Longfin Smelt", "Night Smelt", "Eulachon", "Whitebait Smelt", "Wakasagi", "Surf Smelt", "Delta Smelt" ))
v2$data<-v2$data %>% select(-Species) %>% left_join(v2ddf)
```

```{r}
v2 +
  geom_point(data=df, aes(x=x, y=y, color=Support), pch=18, cex=4, alpha=1) +
  geom_tippoint(aes(fill=Species), pch=21, alpha=1, cex=4, offset=0.01) +
  geom_treescale() +
  xlim(0, 0.1) +
  scale_fill_viridis_d(option = "H") +
  scale_color_manual(values=c("grey50","grey","black"))

ggsave("outputs/104/lfs-with-refs-tree.pdf", width=8.5, height=11)

```

## LFS only


(1) new LFS samples

```{r}
newseq2<-newseq %>% mutate(Location=ifelse(label %in% mynooks$label, "Nooksack","Larval Smelt")) %>% select(-Sort, -Species) %>%
  rename(Sample=label, Reads=Dedup)
newseq2
```

(2) Previous

```{r}
meta140<-read_csv("meta/meta140.csv") %>% select(-Loc, -Number, -File, -Alignment)
meta140
```


```{r}
comb281<-bind_rows(newseq2, meta140)
comb281

write_tsv(select(comb281, Path),"bamlists/281.bamlist",col_names=FALSE)
```
```{sh, eval=FALSE}
srun -p high -t 12:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 36  \
-bam bamlists/281.bamlist -rf $HOME/delta-smelt/metadata/large-contigs.txt \
-ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-minInd 71 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-3 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/104/281-ibs > outputs/104/281-ibs.stdout 2> outputs/104/281-ibs.stderr &


srun -p high -t 12:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 36  \
-bam bamlists/281.bamlist -r lg01 \
-ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-minInd 71 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-3 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/104/281-lg01-ibs > outputs/104/281-lg01-ibs.stdout 2> outputs/104/281-lg01-ibs.stderr &


#Altering minInd/MAF
srun -p med -t 2:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 36  \
-bam bamlists/281.bamlist -r lg01 \
-ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-minInd 141 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-3 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/104/281-lg01-ibs-50 > outputs/104/281-lg01-ibs.stdout 2> outputs/104/281-lg01-ibs.stderr &

srun -p med -t 2:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 36  \
-bam bamlists/281.bamlist -r lg01 \
-ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-minInd 225 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-3 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.05 \
-out outputs/104/281-lg01-ibs-80 > outputs/104/281-lg01-ibs-80.stdout 2> outputs/104/281-lg01-ibs-80.stderr &

#Altering minInd/MAF
srun -p med -t 2:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 36  \
-bam bamlists/281.bamlist -r lg01 \
-ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-minInd 141 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-3 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.1 \
-out outputs/104/281-lg01-ibs-50-1 > outputs/104/281-lg01-ibs.stdout 2> outputs/104/281-lg01-ibs.stderr &

srun -p med -t 2:00:00 --mem=32G --nodes=1 $HOME/angsd/angsd -P 36  \
-bam bamlists/281.bamlist -r lg01 \
-ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-minInd 225 -minMapQ 10 -minQ 20 -GL 2 -doGLF 2 \
-doMajorMinor 1 -doMaf 1 -SNP_pval 1e-3 \
-doIBS 1 -doCounts 1 -doCov 1 -makeMatrix 1 -minMaf 0.1 \
-out outputs/104/281-lg01-ibs-80-1 > outputs/104/281-lg01-ibs-80.stdout 2> outputs/104/281-lg01-ibs-80.stderr &

```

lg01 has 6129 snps
212901 snps total with 1/4 missing

lg01 @ 50% missing MAF 0.05 3858 SNPs
lg01 @ 80% missing MAF 0.05 2218 SNPs

lg01 @ 50% missing MAF 0.10 2419 SNPs     
lg01 @ 80% missing MAF 0.10 1468 SNPs    
```{r}
m <- as.matrix(read.table("outputs/104/281-lg01-ibs-80-1.covMat"))
meta<-comb281
eig <- eigen(m)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
covs<-eig$vectors[,1:3] %>% as_tibble() %>% bind_cols(meta)

text12<-covs %>% select(Sample, Location, V1, V2) %>%
  group_by(Location) %>% summarize(Count=n(), x=mean(V1), y=mean(V2))

covs12<-ggplot(covs) +
  geom_point(aes(x=V1, y=V2, fill=Location), pch=21, alpha=0.75) +
  geom_text_repel(data=text12, aes(x=x, y=y, label=Location), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = "")) +
  theme_bw() +
  theme(legend.position="")


text13<-covs %>% select(Sample, Location, V1, V3) %>%
 group_by(Location) %>% summarize(Count=n(), x=mean(V1), y=mean(V3))

covs13<-ggplot(covs) +
  geom_point(aes(x=V1, y=V3, fill=Location), pch=21, alpha=0.75) +
  geom_text_repel(data=text13, aes(x=x, y=y, label=Location), max.overlaps = Inf) +
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = "")) +
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = "")) +
  theme_bw() +
  theme(legend.position="")


ggarrange(covs12, covs13)
```

     V1      V2    V3 Sample    Reads Path                                                Location
   <dbl>   <dbl> <dbl> <chr>     <dbl> <chr>                                               <chr>   
1 0.0528 -0.0661 0.658 SFBY_075 684288 /home/maccamp/data/longfin/SFBY_075_R1.sort.flt.bam SFBY    
2 0.0499 -0.0561 0.488 SFBY_083 494349 /home/maccamp/data/longfin/SFBY_083_R1.sort.flt.bam SFBY    

```{r}
covs12
ggsave("outputs/104/first-sequencing-run-pcs.jpeg")
```
```{sh, eval=FALSE}
srun -p high -t 00:30:00 --nodes=1 python $HOME/pcangsd/pcangsd.py -beagle outputs/104/initial-lg01.beagle.gz \
    -o outputs/102/initial-pca -threads 10
```

```{sh, eval=FALSE}
for i in {1..12};
do

    for j in {1..1};
    do echo -n "$i - $j ";
    srun -p high -t 4:00:00 $HOME/angsd/misc/NGSadmix -likes outputs/108/239-ibs.beagle.gz -K $i -o outputs/108/ngsadmix239/ngsadmix-k$i-run$j -P 36;
    done;

  echo "";
done;
```