---
title: "106-new-seq"
output: html_document
date: "2023-04-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ggrepel)
library(ggpubr)
```

```{r}
meta<-read_csv("meta/meta260.csv")
meta$Location<-factor(meta$Location, levels=c("YBAK","SKNA","PTLC","HRLC","FRAS","LWSH",
                                              "NOOK","COLR",
                                              "HUMB","PETA","ALVS","SUIB","CHPI",
                                         "New Sequencing"))

meta
```
## Have seqs

Will try to control for effects a bit. Making a region file from 101 to restrict snps to previously discovered ones. Using trimmed and downsampled sequences from 105.

(base) maccamp@farm:~/longfin/outputs/101$ gunzip -c sub140-contigs.beagle.gz | cut -f 1 | perl -pe 's/_/:/g' > ../106/regions.txt
grep lg01 regions.txt > lg01.txt
grep ^lg regions.txt > lgs.txt
60639 sites

```{eval=FALSE}
srun -p high -t 14:00:00 --mem=16G --nodes=1 $HOME/angsd/angsd -P 8 \
  -ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
  -bam $HOME/longfin/bamlists/260.bamlist -GL 1 \
  -doGLF 2 -doMajorMinor 1 -doMaf 2 -SNP_pval 1e-6 -minMapQ 10 -minQ 20 -minMaf 0.05 \
  -rf $HOME/longfin/outputs/106/lg01.txt -minInd 208 \
  -out $HOME/longfin/outputs/106/260-txt > outputs/106/beagle.out 2> outputs/106/beagle.err &

  #1674 snps
  
srun -p high -t 14:00:00 --mem=16G --nodes=1 $HOME/angsd/angsd -P 8 \
  -ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
  -bam $HOME/longfin/bamlists/260.bamlist -GL 1 \
  -doGLF 2 -doMajorMinor 1 -doMaf 2 -SNP_pval 1e-6 -minMapQ 10 -minQ 20 \
  -rf $HOME/longfin/outputs/106/lg01.txt -minInd 208 \
  -out $HOME/longfin/outputs/106/260-no-maf > outputs/106/beagle-maf.out 2> outputs/106/beagle-maf.err &  
  
python $HOME/pcangsd/pcangsd.py -beagle outputs/106/260-txt.beagle.gz -o outputs/106/260-lg01 -threads 10
#1674 sites without MAF filtercs 

srun -p high -t 32:00:00 --mem=16G --nodes=1 $HOME/angsd/angsd -P 8 \
  -ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
  -bam $HOME/longfin/bamlists/260.bamlist -GL 1 \
  -doGLF 2 -doMajorMinor 1 -doMaf 2 -SNP_pval 1e-6 -minMapQ 10 -minQ 20 -minMaf 0.05 \
  -rf $HOME/longfin/outputs/106/lgs.txt -minInd 208 \
  -out $HOME/longfin/outputs/106/260-lgs > outputs/106/beagle-lgs.out 2> outputs/106/beagle-lgs.err &

#51259 sites
  
conda activate py2
srun -p high -t 01:00:00 python $HOME/pcangsd/pcangsd.py -beagle outputs/106/260-lgs.beagle.gz \
  -o outputs/106/260-lgs -threads 10 
conda deactivate

```


```{r}
#' @param samples character vector with the individuals IDs in the order in which
#' they were passed in the bamlist to angsd.
#' @param cov covariance matrix
covar2pcs <- function(samples, cov) {
  
  
  eig <- eigen(cov, symm = TRUE)
  PC <- as.data.frame(eig$vectors) %>%
    as_tibble() %>%
    setNames(sprintf("PC-%02d", 1:ncol(.)))
  
  samtib <- tibble(sample = samples)
  
  list(
    PCs = bind_cols(samtib, PC),
    eigevalues = eig$values
  )
}
```

```{r}
#cov<-read_delim("outputs/106/260-lg01.cov", col_names=FALSE, delim=" ") %>% as.matrix()
cov<-read_delim("outputs/106/260-lgs.cov", col_names=FALSE, delim=" ") %>% as.matrix()
```

```{r}
pca <- covar2pcs(meta$Sample, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

pp_meta <- pca_pairs %>%   # just keep the first 6 PCs around
  left_join(., meta, by = c("sample" = "Sample")) 
```


```{r}
npc <- 3
pp_meta2 <- pp_meta %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) ) %>%
  mutate(Origin=ifelse(Location=="New Sequencing","New","Old"))

eig <- eigen(cov, symm = TRUE)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
```

```{r}
ggplot(pp_meta2, aes(x = val_x, y = val_y)) +
  geom_point(aes(color=Location)) + 
  facet_grid(PCx ~ PCy)
```

```{r}
sub12<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-02") )
text<- sub12 %>% select(sample,Location,val_x,val_y) %>% group_by(Location) %>%
  mutate(meanX=mean(val_x), meanY=mean(val_y)) %>% select(Location, meanX, meanY) %>%
  unique()   

pc12<-ggplot(sub12, aes(x = val_x, y = val_y, color=Location))+
  geom_point(size = 2, alpha=0.75) +
  geom_label_repel(data=text, aes(x=meanX, y=meanY, color=Location,
                                  label=Location), max.overlaps = Inf) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = ""))+
  theme(legend.position = "")+
  theme(plot.title = element_text(hjust = 0.5)) 

sub13<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-03") )
text13<- sub13 %>% select(sample,Location,val_x,val_y) %>% group_by(Location) %>%
  mutate(meanX=mean(val_x), meanY=mean(val_y)) %>% select(Location, meanX, meanY) %>%
  unique()   

pc13<-ggplot(sub13, aes(x = val_x, y = val_y, color = Location)) +
  geom_point(size = 2, alpha=0.75) +
  geom_label_repel(data=text13, aes(x=meanX, y=meanY, color=Location,
                                  label=Location), max.overlaps = Inf) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = ""))+
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")

pcs<-ggarrange(pc12, pc13, ncol= 2, widths=c(1,1))
pcs
ggsave("outputs/106/combined-pca-260.jpeg", width=10)
```

```{sh, eval=FALSE}
srun -p high -t 01:00:00 --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/106/260-lgs.beagle.beagle.gz -K 2 -o outputs/106/260-lgs.K2
srun -p high -t 01:00:00 --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/106/260-lgsb.eagle.beagle.gz -K 3 -o outputs/106/260-lgs.K3
srun -p     high -t 01:00:00 --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/106/260-lgs.beagle.beagle.gz -K 4 -o outputs/106/260-lgs.K4
srun -p high -t 01:00:00 --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/106/260-lgs.beagle.beagle.gz -K 5 -o outputs/106/260-lgs.K5

```

## Phylogeny

Rooting....? Let's try rooting with night smelt 

```{r}

night<-read_csv("meta/non-lfs-meta.csv") %>% filter(Species %in% c("Whitebait Smelt","Night Smelt")) %>% rename(Loc=Species) %>% mutate(Location=Loc) %>% select(Sample, Location, Dedup, Path, Loc) %>% filter(Dedup > 6e5)

phylo<-bind_rows(meta,night)
write_tsv(phylo %>% select(Path), col_names =FALSE, file="bamlists/phylo271.bamlist")
```

Gives us 11 fish. 271 total

```{sh, eval=FALSE}
srun -p high -t 38:00:00 --mem=16G --nodes=1 --ntasks=1 angsd -P 12 \
-bam bamlists/phylo271.bamlist \
-out /home/maccamp/longfin/outputs/106/plink \
-anc /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
-rf $HOME/longfin/outputs/106/lgs.txt  \
-minInd 244 -minMaf 0.05  -minMapQ 10 -minQ 20 -GL 1 -doMajorMinor 1 -doMaf 1 -SNP_pval 1e-6 \
-doGeno 4 -doPost 1 -postCutoff 0.95 -doPlink 2 >outputs/106/plink.out 2>outputs/106/plink.stderr &

#next, can restrict to certain sites
```


## With NOOK only

```{sh, eval=FALSE}
srun -p med -t 1-12:00:00 --mem=16G --nodes=1 $HOME/angsd/angsd -P 22 \
  -ref /home/maccamp/genomes/hypomesus-20210204/Hyp_tra_F_20210204.fa \
  -bam $HOME/longfin/bamlists/bamlist134.bamlist -GL 1 \
  -doGLF 2 -doMajorMinor 1 -doMaf 2 -SNP_pval 1e-6 -minMapQ 10 -minQ 20 -minMaf 0.05 \
  -rf $HOME/longfin/outputs/106/lgs.txt -minInd 117 \
  -out $HOME/longfin/outputs/106/134-lgs > outputs/106/134-beagle-lgs.out 2> outputs/106/134-beagle-lgs.err &

#conda clean --all 
#conda create -n py2 python=2.7 scipy numpy pandas

#conda activate python2
# actual used pip to install scipy

#Both worked..
#42735 sites
python $HOME/pcangsd/pcangsd.py -beagle outputs/106/134-lgs.beagle.gz -o outputs/106/134-lgs -threads 10

```



```{r}
cov<-read_delim("outputs/106/134-lgs.cov", col_names=FALSE, delim=" ") %>% as.matrix()
m134<-meta %>% filter(Location != "New Sequencing")
```

```{r}
pca <- covar2pcs(m134$Sample, cov)

pca_long <- pca$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expg <- expand.grid(sample = pca$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairs <- dplyr::left_join(expg, pca_long, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_long, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

pp_meta <- pca_pairs %>%   # just keep the first 6 PCs around
  left_join(., m134, by = c("sample" = "Sample")) 
```


```{r}
npc <- 3
pp_meta2 <- pp_meta %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) ) %>%
  mutate(Origin=ifelse(Location=="New Sequencing","New","Old"))

eig <- eigen(cov, symm = TRUE)
var<-eig$values/sum(eig$values)
cumvar<-cumsum(eig$values)/sum(eig$values)

head(var)
head(cumvar)
```

```{r}
ggplot(pp_meta2, aes(x = val_x, y = val_y)) +
  geom_point(aes(color=Location)) + 
  facet_grid(PCx ~ PCy)
```

```{r}
sub12<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-02") )
text<- sub12 %>% select(sample,Location,val_x,val_y) %>% group_by(Location) %>%
  mutate(meanX=mean(val_x), meanY=mean(val_y)) %>% select(Location, meanX, meanY) %>%
  unique()   

pc12<-ggplot(sub12, aes(x = val_x, y = val_y, color=Location))+
  geom_point(size = 2, alpha=0.75) +
  geom_label_repel(data=text, aes(x=meanX, y=meanY, color=Location,
                                  label=Location), max.overlaps = Inf) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC2", " ", round((100*var[2]),2), "%", sep = ""))+
  theme(legend.position = "")+
  theme(plot.title = element_text(hjust = 0.5)) 

sub13<-pp_meta2 %>% filter( (PCx =="PC-01") & (PCy =="PC-03") )
text13<- sub13 %>% select(sample,Location,val_x,val_y) %>% group_by(Location) %>%
  mutate(meanX=mean(val_x), meanY=mean(val_y)) %>% select(Location, meanX, meanY) %>%
  unique()   

pc13<-ggplot(sub13, aes(x = val_x, y = val_y, color = Location)) +
  geom_point(size = 2, alpha=0.75) +
  geom_label_repel(data=text13, aes(x=meanX, y=meanY, color=Location,
                                  label=Location), max.overlaps = Inf) +
  scale_fill_discrete(na.value = "white") + 
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*var[1]),2), "%", sep = ""))+
  ylab(paste("PC3", " ", round((100*var[3]),2), "%", sep = ""))+
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")

pcs<-ggarrange(pc12, pc13, ncol= 2, widths=c(1,1))
pcs
ggsave("outputs/106/combined-pca-134.jpeg", width=10)
```


```{sh, eval=FALSE}
srun -p high -t 01:00:00 --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/106/134-lgs.beagle.gz -K 2 -o outputs/106/134-lgs.K2
srun -p high -t 01:00:00 --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/106/134-lgs.beagle.gz -K 3 -o outputs/106/134-lgs.K3 &
srun -p     high -t 01:00:00 --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/106/134-lgs.beagle.gz -K 4 -o outputs/106/134-lgs.K4 &
srun -p high -t 01:00:00 --nodes=1 $HOME/angsd/misc/NGSadmix -P 12 -likes outputs/106/134-lgs.beagle.gz -K 5 -o outputs/106/134-lgs.K5 &
```


